\textbf{Automated Reasoning\\Practical Assignment -- Part 1}

### 2022-09-19 docs/paper.tex
Add description of chips implementation

\documentclass[a4paper]{article}

@@

\author{Jelmer Firet, Bram Pulles}
\title{\textbf{Automated Reasoning\\Practical Assignment -- Part 1}}

@@

>>>>

\documentclass{scrartcl}
\usepackage{amsmath}
\usepackage{fancyhdr}
\usepackage{minted}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{svg}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{scalerel}

@@

\author{Jelmer Firet (s1023433, \href{mailto:jelmer.firet@ru.nl}{jelmer.firet@ru.nl}) \and
        Bram Pulles()}
\title{\textbf{Automated Reasoning\\Practical Assignment -- Part 1}}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{Automated Reasoning PA 1|Jelmer Firet \& Bram Pulles}
\fancyhead[R]{\thepage}

\setminted{autogobble,tabsize=4,fontsize=\scriptsize}
\setlength{\parindent}{0pt}

@@

We consider the chip as a $30\times30$ grid. We number the rows from top to bottom, so the top row has index $0$ and the bottom row has index $29$. The columns are numbered the same way from left tot right.

\subsection{variables} % (fold)
\label{sub:chip_variables}
	For each component $i$ we introduce $4$ integer variables: $\mathtt{left}_i$, $\mathtt{right}_i$, $\mathtt{top}_i$ and $\mathtt{bottom}_i$.\\
	These variables indicate where the edges of component $i$ are.\\

	\begin{tabular}{ll}
		$\mathtt{left}_i$ & The left-most column that contains component $i$\\
		$\mathtt{right}_i$ & The left-most column to the right of component $i$\\
		$\mathtt{top}_i$ & The top-most column that contains component $i$\\
		$\mathtt{bottom}_i$ & The top-most column below component $i$\\
	\end{tabular}\\

	Note that $\mathtt{left}_i$ is inclusive; the corresponding column contains component $i$.\\
	However, $\mathtt{right}_i$ is exclusive; the corresponding column does not contain component $i$.\\
	Inclusive-exclusive indexing ensures that $\mathtt{right}_i-\mathtt{left}_i$ is the width of component $i$.\\
	We also use this indexing for the rows, so $\mathtt{bottom}_i-\mathtt{top}_i$ is the height of component $i$.\\
	Component $s$ consists of the cells $(r,c)$ with $\mathtt{left}_i \leq c < \mathtt{right}_i$ and $\mathtt{top}_i \leq r < \mathtt{bottom}_i$

	\begin{figure}[H]
		\centering
		\begin{subfigure}[b]{0.30\textwidth}
			\centering
			\includesvg[width=\textwidth]{chips-indexes}
		\end{subfigure}
		\begin{subfigure}[b]{0.5\textwidth}
			\centering
			\begin{align*}
				\mathtt{left}_i&=1 & \mathtt{right}_i &= 4\\
				\mathtt{top}_i&=1 & \mathtt{bottom}_i &= 3
			\end{align*}
			\begin{align*}
				\mathtt{width}_i=\mathtt{right}_i-\mathtt{left}_i&=4-1=3\\
				\mathtt{height}_i=\mathtt{bottom}_i-\mathtt{top}_i&=3-1=2
			\end{align*}
		\end{subfigure}
	\end{figure}
% subsection chip_variables (end)

\subsection{Constraints} % (fold)
\label{sub:chip_constraints}
\begin{enumerate}
	\item 
	We want every component to be placed inside the $30\times 30$ grid.\\
	For this we add four constrains for each component:
	\begin{align*}
		0 &\leq \mathtt{left}_i & \mathtt{right}_i &\leq 30 & 0 &\leq \mathtt{top}_i & \mathtt{bottom}_i &\leq 30
	\end{align*}
	For every cell $(r,c)$ of component $i$ we then have:
	\begin{align*}
		0 \leq \mathtt{left}_i \leq c <  \mathtt{right}_i \leq 30 &&
		0 \leq \mathtt{top}_i \leq r < \mathtt{bottom}_i \leq 30
	\end{align*}
	So each cell of the component is within the grid.

	\item 
	We want each component to have the correct size.\\
	Suppose component $i$ should have size $(w,h)$ then we add the constraint:
	\begin{align*}
		(\mathtt{right}_i-\mathtt{left}_i = w\;&\land\; \mathtt{bottom}_i-\mathtt{top}_i = h)\;\lor\\
		(\mathtt{right}_i-\mathtt{left}_i = h\;&\land\; \mathtt{bottom}_i-\mathtt{top}_i = w)
	\end{align*}
	The first line says that the component has the right size when it is not rotated.\\
	The second line says that the component has the right size if it is rotated 90$^\circ$.\\
	Because we want the component to have the right size in either orientation, we combine these clauses with an $\lor$.

	\item 
	We do not want any of the components to overlap.
	For every pair of components $i<j$ we add the following constraint:
	\begin{align*}
		\mathtt{right}_i &\leq \mathtt{left}_j &\lor&&
		\mathtt{left}_i &\geq \mathtt{right}_j &\lor&&
		\mathtt{bottom}_i &\leq \mathtt{top}_j &\lor&&
		\mathtt{top}_i &\geq \mathtt{bottom}_j
	\end{align*}
	This constrains say that component $j$ is entirely to the right, to the left, below or above component $i$. In each of these cases there is no overlap possible.

	\item 
	We want every regular component to share part of an edge with a power component.\\
	For every regular component $i$ we add the following constraint:
	\begin{align*}
		\scalerel*{\lor}{\sum}_{p\in P}\left(
		\begin{array}{l}
			(\mathtt{right}_i = \mathtt{left}_p \;\land\; \mathtt{top}_i < \mathtt{bottom}_p \;\land\; \mathtt{top}_p < \mathtt{bottom}_i)\;\lor\\
			(\mathtt{left}_i = \mathtt{right}_p \;\land\; \mathtt{top}_i < \mathtt{bottom}_p \;\land\; \mathtt{top}_p < \mathtt{bottom}_i)\;\lor\\
			(\mathtt{bottom}_i = \mathtt{top}_p \;\land\; \mathtt{left}_i < \mathtt{right}_p \;\land\; \mathtt{left}_p > \mathtt{right}_i)\;\lor\\
			(\mathtt{top}_i = \mathtt{bottom}_p \;\land\; \mathtt{left}_i < \mathtt{right}_p \;\land\; \mathtt{left}_p > \mathtt{right}_i)
		\end{array}
		\right)
	\end{align*}

	This statement starts with a big or over the power components. This ensures that there is a power component $p$ that touches component $i$.\\
	Within the parenthesis we find four cases: component $i$ touches the left, the right, the top or the bottom of power component $p$.
	To check if two components touch we need two parts.\\
	First, both components need to touch a vertical or horizontal line from different sides. This is ensured by the first condition of every case.\\
	Second, they need to share a part of that line, which is ensured by the other conditions.
	We will only explain these other conditions for the first case.

	In the first case the components touch a vertical line.
	For the components to share a part of this line, there must be a row $r$ that contains both components.\\
	This line must satisfy $\mathtt{top}_i \leq r < \mathtt{bottom}_i$ and $\mathtt{top}_p \leq r < \mathtt{bottom}_p$\\
	We can combine these inequalities to form the following four conditions:
	\begin{align*}
		\mathtt{top}_i &< \mathtt{bottom}_i &\mathtt{top}_i &< \mathtt{bottom}_p & \mathtt{top}_p &< \mathtt{bottom}_i &\mathtt{top}_p &< \mathtt{bottom}_p
	\end{align*}
	The first and last condition are always satisfied (by part 2 and $0<w,h$).\\
	The other conditions are added to the first case to guarantee overlap.

	\item The centers of the power components are far enough apart
	For every pair $p < q$ of power components we add the following constraint:
	\begin{align*}
		((\mathtt{top}_p + \mathtt{bottom}_p) - (\mathtt{top}_q + \mathtt{bottom}_q) &\geq 2\cdot D)\;\lor\\
		((\mathtt{top}_q + \mathtt{bottom}_q) - (\mathtt{top}_p + \mathtt{bottom}_p) &\geq 2\cdot D)\;\lor\\
		((\mathtt{left}_p + \mathtt{right}_p) - (\mathtt{left}_q + \mathtt{right}_q) &\geq 2\cdot D)\;\lor\\
		((\mathtt{left}_q + \mathtt{right}_q) - (\mathtt{left}_p + \mathtt{right}_p) &\geq 2\cdot D)
	\end{align*}

	There are four cases: the center of component $p$ is $D$ units below, above, to the left or to the right of the center of component $q$. To ensure the vertical distance between the centers of two components we could use the following formula:
	\begin{align*}
		\lvert(\mathtt{top}_p + \mathtt{bottom}_p)/2 - (\mathtt{top}_q + \mathtt{bottom}_q)/2\rvert \geq D
	\end{align*}
	However we would like a formula without division or absolute values.
	To get rid of the division we multiply everything by $2$.
	We remove the absolute value by splitting the equation into two cases.
	This equation then becomes the first two cases of our constraint.
	The other two cases are found by considering the horizontal distance.
\end{enumerate}
% subsection chip_constraints (end)

\subsection{Result} % (fold)
\label{sub:result}
The largest $D$ for which we can find a result is $17.5$.
For $D=17.5$ it finds a result in $1.865$ seconds.
For $D=18$ it proves that it is unsatisfiable in $67.782$ seconds.

A solution for $D=17.5$ is illustrated below:
\begin{figure}[H]
	\centering
	\includesvg[width=0.7\textwidth]{chips-solution}
\end{figure}

% subsection result (end)
